---
title: "Logit Model - NCAA (Qyalys)"
author: John Antony
date: March 2nd, 2017
output:
  slidy_presentation: default
  ioslides_presentation: default
  beamer_presentation: default
---


### Logit Example

For the NCAA data, take the top 32 teams and make their dependent variable 1, and that of the bottom 32 teams zero. Therefore, the teams that have $y=1$ are those that did not lose in the first round of the playoffs, and the teams that have $y=0$ are those that did. Estimation is done by maximizing the log-likelihood.  


```{r}
#Importing and displaying the NCAA dataset
ncaa = read.table("data_files/ncaa.txt",header=TRUE)    

y1 = 1:32
y1 = y1*0+1
y2 = y1*0
y = c(y1,y2)
```
NCAA Team Data
```{r}
ncaa
```

----
```{r}
#Running the binomial logit function on the x variables for regression
x = as.matrix(ncaa[4:14])
h = glm(y~x, family=binomial(link="logit"))
names(h)
```
----
```{r}
print(logLik(h))
summary(h)
```
```{r}
print(h)
```

----
Printing the probability of each team winning the next match in the playoff. This model can be used for running probability of a sale (lead scoring) using various marketing/sales attributes. 

```{r}
h$fitted.values
```

